Analyzing tweets to detect social media events

### 2020-05-20 Rmd/abstract.Rmd
upd
updates to methods
many updates
rewrote abstract

We develop
a mentoring framework to guide undergraduate researchers
through individualized research projects in social media data science. 
Our framework involves research question formulation, data acquisition, data 
analysis and visualization, and presentation and communication of results. Our two honors students, 
whose projects serve as case studies for our framework, completed all components of the 
individualized research projects. We found that data science research skills, self-confidence in 
research ability, and professional interest in data science increased for both 
students. We describe our successes, lessons learned, and ideas for others to 
build similar frameworks.

>>>>

We describe methods for collecting tweets from Twitter. With these methods, we 
curate a large collection of tweets.
We present methods for social media event detection through text analysis of tweets. 
We describe our use of tweets in mentored undergraduate research. 
We also offer other directions for other text analysis of tweets.

### 2020-05-20 Rmd/jse-2019-ms.Rmd
rewrote abstract

While our tweet collection enables us to address many possible research questions, the 
dynamic content of tweets over time particularly piqued our interest. Together, students and 
mentors hypothesized that high-profile social media events would generate a high volume of 
tweets, and that we'd detect social media events through changes in tweet 
topic content over time. We present below 1) an approach for collecting tweets in real 
time and 2) statistical methods for detecting social media 
events via latent Dirichlet allocation modeling of collections of tweets and 3) reflections on 
using this data set in research mentoring of undergraduate students.

>>>>

While our tweet collection enables us to address many possible research questions, the 
dynamic content of tweets over time particularly piqued our interest. Together, students and 
mentors hypothesized that high-profile social media events would generate a high volume of 
tweets, and that we would detect social media events through changes in tweet 
topic content over time. We present below 1) an approach for collecting tweets in real 
time and 2) statistical methods for detecting social media 
events via latent Dirichlet allocation modeling of collections of tweets and 3) reflections on 
using this data set in research mentoring of undergraduate students.
